\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-08 Documentation/TearingModeTheory.tex
Additions to tearing mode writeup

In a tokamak plasma,  the ambient magnetic field, 
${\bf B}$,  is strong enough to significantly alter charged particle
trajectories. Consequently,
tokamak plasmas are highly  anisotropic, responding differently to
forces that are parallel and perpendicular to the direction
of ${\bf B}$. As is well known, charged particles respond to the Lorentz force
by freely streaming in the direction of ${\bf B}$, while executing
circular  {\em gyro-orbits} in the plane perpendicular to ${\bf B}$ (Fitzpatrick 2008). 
The typical   {\em gyroradius}\/ of a charged particle 
gyrating in a
magnetic field is given by
\begin{equation}
\rho_s = \frac{v_{t\,s}}{{\mit\Omega}_s},
\end{equation}
where
\begin{equation}
{\mit\Omega}_s = \frac{e\,B}{m_s}
\end{equation}
is the  {\em gyrofrequency}\/ associated with the
gyration.

@@

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Wesson 1978; Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}\simeq 0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(By definition, $a\ll R_0$ in a large aspect-ratio tokamak. In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}\simeq 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

\subsection{Solution in Presence of Resistive Wall}\label{resistive}
Suppose, now, that the wall at $r=r_w$ possesses a finite electrical resistivity, but is surrounded by a second perfectly-conducting wall
located at radius $r_c>r_w$.
The solution in the outer region can be written
\begin{equation}\label{e111}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is specified in the previous section, and the real function  $\hat{\psi}_w(r)$ is a solution of 
\begin{equation}\label{e100a}
\frac{d^2\hat{\psi}_w}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_w}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_w - \frac{J_z'\,\hat{\psi}_w}{r\,(1/q-n/m)}\simeq 0
\end{equation}
that satisfies
\begin{align}\label{e101a}
\hat{\psi}_w(r\leq r_s) &= 0,\\[0.5ex]
\hat{\psi}_w(r_w) &= 1,\label{e102a}\\[0.5ex]
\hat{\psi}_w(r\geq r_c) &= 0.\label{e103a}
\end{align}
It is easily seen that
\begin{equation}\label{e117}
\hat{\psi}_w(r_w< r < r_c) = \frac{(r/r_c)^{-m} - (r/r_c)^m}{(r_w/r_c)^{-m} - (r_w/r_c)^m}.
\end{equation}

@@

The solution in the outer region can be written
\begin{equation}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r) + {\mit\Psi}_c\,\hat{\psi}_c(r),
\end{equation}
where the real functions $\hat{\psi}_s(r)$ and $\hat{\psi}_w(r)$ are specified in Sections~\ref{perfect} and \ref{resistive}, respectively. 
Moreover, 
the real function  $\hat{\psi}_c(r)$ is a solution of 
\begin{equation}\label{e100b}
\frac{d^2\hat{\psi}_c}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_c}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_c\simeq 0
\end{equation}
that satisfies
\begin{align}
\hat{\psi}_c(r\leq r_w) &= 0,\\[0.5ex]
\hat{\psi}_c(r_c) &= 1,\\[0.5ex]
\hat{\psi}_c(\infty) &= 0.
\end{align}
It is easily seen that
\begin{align}
\hat{\psi}_c(r\leq r_w) &= 0,\\[0.5ex]
\hat{\psi}_c(r_w<r \leq r_c) &= \frac{(r/r_w)^m-(r/r_w)^{-m}}{(r_c/r_w)^m - (r_c/r_w)^{-m}},\label{e142}\\[0.5ex]
\hat{\psi}_c(r> r_c) &=\left(\frac{r}{r_c}\right)^{-m}.\label{e143}
\end{align}

@@

Taking the scalar product of  $R_0\,{\bf e}_z$ with the plasma equation of motion, (\ref{e175}), and integrating around magnetic
flux-surfaces, similar arguments to those just employed for the ion poloidal flow allow us to write the zeroth-order ion toroidal angular velocity profile in the form
\begin{equation}
{\mit\Omega}_\varphi(r,t)= {\mit\Omega}_{\varphi\,0}(r) + {\mit\Delta\Omega}_\varphi(r,t),
\end{equation}
where ${\mit\Omega}_{\varphi\,0}(r)$ is the 
unperturbed (by the tearing mode) ion toroidal angular velocity profile, and ${\mit\Delta\Omega}_{\theta}(r,t)$ is the  modification to  this  profile
induced by the toroidal electromagnetic torque that develops at the resonant surface. The perturbed ion toroidal angular
equation of motion is written
\begin{equation}\label{e185}
4\pi^2\,R_0^{\,3}\left[\rho\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s).
\end{equation}
Note that the ion parallel viscosity tensor does not give rise to damping of the toroidal flow profile. Furthermore, there is  no
neoclassical enhancement of the plasma toroidal ion inertia.

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In general, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_s\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_s\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_s=\mu_\perp(r_s)$.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf b}\right] \simeq \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel$ and $\eta_\perp=1/\sigma_\perp$ are the parallel and perpendicular plasma
resistivities,  respectively. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$.

Next, we can add the electron and ion equations of motion, (\ref{e3.79b}) and (\ref{e3.80b}), again neglecting electron inertia and electron viscosity, 
to give the {\em plasma equation of motion}: 
\begin{equation}
n_0\,m_i\!\left[\frac{\partial {\bf V}}{\partial t} + ({\bf V}\cdot\nabla ){\bf V} + \frac{1}{1+\tau}\,({\bf V}_\ast\cdot\nabla){\bf V}_E\right]
+ \nabla p + \nabla\cdot\bpi_i - {\bf j}\times {\bf B}\simeq 0.
\end{equation}
Here, $\nabla\cdot\bpi_i$ includes the contribution from the anomalous (i.e., enhanced due to the action of small-scale plasma turbulence) perpendicular viscosity tensor. The contribution from the gyroviscosity tensor,
which is the same size as the ion inertial terms, has already been incorporated into the equation (Hazeltine \& Meiss 1992). For the moment, we are neglecting the contribution of the parallel viscosity tensor.

Finally, we can combine the electron and ion energy evolution equations, (\ref{e3.79c}) and (\ref{e3.80c}), to obtain
a {\em plasma energy evolution equation}:
\begin{equation}
\frac{\partial p}{\partial t} + {\bf V}\cdot\nabla p + {\mit\Gamma}\,p\,\nabla\cdot{\bf V} + \frac{2}{3}\,\nabla\cdot {\bf q}\simeq 0, 
\label{e223}
\end{equation}
where ${\mit\Gamma}=5/3$. Here, we have neglected a number of  terms (e.g., the ohmic heating terms) because they
are, at least, factors $({\mit\Omega}_e\,\tau_e)^{-1}$ or  $({\mit\Omega}_i\,\tau_i)^{-1}$ smaller than $\nabla\cdot{\bf q}$. We have also neglected the viscous heating term because we
expect the dominant contribution, which comes from the parallel viscosity, to be much smaller than its nominal magnitude in a large aspect-ratio tokamak
plasma.

@@

It is convenient to adopt the following normalization scheme that renders all quantities in the drift-MHD equations dimensionless: 
$\hat{\nabla} = l\,\nabla$, $\hat{d}_i=d_i/l$, $\hat{t} = t/(l/V_A)$, $\hat{\bf B} = {\bf B}/B_z$, $\hat{\bf E} = {\bf E}/(B_z\,V_A)$,
$\hat{\bf j} = {\bf j}/(B_z/\mu_0\,l)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,V_A\,l)$, $\hat{\bf V} = {\bf V}/V_A$,
$\hat{{\bf V}}_{\ast,i}= {\bf V}_{\ast,i}/V_A$, $\hat{V}_{\parallel\,i}= V_{\parallel\,i}/V_A$, $\hat{p} = p/(B_z^{\,2}/\mu_0)$, $\hat{p}_0 = p_0/(B_z^{\,2}/\mu_0)$, $\hat{\bpi}_{i} = \bpi_{i}/(B_z^{\,2}/\mu_0)$,
$\hat{\bf q} = {\bf q} /(B_z^{\,2}\,V_A/\mu_0)$. Equations~(\ref{e220})--(\ref{e226}) yield
\begin{align}\label{e227}
\hat{\bf E} + \hat{\bf V}\times \hat{\bf B} + \hat{d}_i\!\left[\hat{\nabla} \hat{p} + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\hat{\nabla} \hat{p})\,{\bf b} + \hat{\nabla}\cdot\hat{\bpi}_e-\hat{\bf j}\times\hat{\bf b}\right] &\simeq \hat{\eta}_\parallel\,\hat{\bf j}_\parallel+ \hat{\eta}_\perp\,\hat{\bf j}_\perp,\\[0.5ex]
\frac{\partial \hat{\bf V}}{\partial \hat{t}} 
+ (\hat{\bf V}\cdot\hat{\nabla} )\hat{\bf V} + 
\frac{1}{1+\tau}\,(\hat{\bf V}_\ast\cdot\hat{\nabla})\hat{\bf V}_E
+ \hat{\nabla} \hat{p} + \hat{\nabla}\cdot\hat{\bpi}_i -
 \hat{\bf j}\times \hat{\bf B} &\simeq 0,
 \\[0.5ex]
\frac{\partial \hat{p}}{\partial \hat{t}} + \hat{\bf V}\cdot\hat{\nabla} \hat{p} + {\mit\Gamma}\,\hat{p}\,\hat{\nabla}\cdot\hat{\bf V} + \frac{2}{3}\,\hat{\nabla}\cdot \hat{\bf q}&\simeq 0,\label{e229}
\end{align}
and
\begin{align}
\hat{\nabla}\cdot\hat{\bf B} &= 0,\label{e230}\\[0.5ex]
\hat{\nabla}\times \hat{\bf E} &= - \frac{\partial \hat{\bf B}}{\partial \hat{t}},\label{e231}\\[0.5ex]
\hat{\bf j} &= \hat{\nabla} \times \hat{\bf B}.\label{e232}
\end{align}
Finally,
\begin{align}
\hat{\bf V}_E &= \hat{\bf E}\times {\bf b},\\[0.5ex]
\hat{\bf V} &= \hat{\bf V}_E + \hat{V}_{\parallel\,i}\,{\bf b},\\[0.5ex]
\hat{\bf V}_\ast &= \hat{d}_i\,{\bf b}\times \hat{\nabla} \hat{p},\\[0.5ex]
\hat{\bf V}_i &= \hat{\bf V} + \frac{\hat{\bf V}_{\ast\,i}}{1+\tau}.
\end{align}

>>>>

In a tokamak plasma,  the ambient magnetic field, 
${\bf B}$,  is strong enough to significantly alter charged particle
trajectories. Consequently,
tokamak plasmas are highly  anisotropic, responding differently to
forces that are parallel and perpendicular to the direction
of ${\bf B}$. As is well known, charged particles react to the Lorentz force
by freely streaming in the direction of ${\bf B}$, while executing
circular  {\em gyro-orbits} in the plane perpendicular to ${\bf B}$ (Fitzpatrick 2008). 
The typical   {\em gyroradius}\/ of a charged particle 
gyrating in a
magnetic field is given by
\begin{equation}
\rho_s = \frac{v_{t\,s}}{{\mit\Omega}_s},
\end{equation}
where
\begin{equation}
{\mit\Omega}_s = \frac{e\,B}{m_s}
\end{equation}
is the  {\em gyrofrequency}\/ associated with the
gyration.

@@

\subsection{Cylindrical Tearing Mode Equation}
To lowest order, the tearing perturbation is governed by the linearized version of Equation~(\ref{e84}), which yields
\begin{equation}\label{e96}
(\delta{\bf B}\cdot \nabla)\,{\bf j} + ({\bf B}\cdot\nabla)\,\delta{\bf j} - (\delta{\bf j}\cdot\nabla)\,{\bf B} - 
({\bf j}\cdot\nabla)\,\delta {\bf B}\simeq {\bf 0}.
\end{equation}
Making use of Equations~(\ref{e90})--(\ref{e94}), the $z$-component of the previous equation gives
the {\em cylindrical tearing mode equation}\/ (Wesson 1978; Fitzpatrick 1993),
\begin{equation}
\frac{\partial^2\psi}{\partial r^2} + \frac{1}{r}\,\frac{\partial\psi}{\partial r}-\frac{m^2}{r^2}\,\psi - \frac{J_z'\,\psi}{r\,(1/q-n/m)}=  0,
\end{equation}
where 
\begin{equation}
J_z(r)= \frac{R_0\,\mu_0\,j_z(r)}{B_z}.
\end{equation}

\subsection{Solution in Presence of Perfectly-Conducting Wall}\label{perfect}
Suppose that the plasma occupies the region $0\leq r\leq a$, where $a$ is the plasma {\em minor radius}. It follows that
$J_z(r)=0$ for $r>a$. Let the plasma be surrounded by a concentric, rigid, radially-thin, perfectly-conducting, wall of radius $r_w>a$. 
(By definition, $a\ll R_0$ in a large aspect-ratio tokamak. In most circumstances, the wall represents the metal vacuum vessel that surrounds  the plasma.) An appropriate
physical solution of the cylindrical tearing mode equation takes the separable form
\begin{equation}\label{e99}
\psi(r,t)= {\mit\Psi}_s(t)\,\hat{\psi}_s(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is a solution of 
\begin{equation}\label{e100}
\frac{d^2\hat{\psi}_s}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_s}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_s - \frac{J_z'\,\hat{\psi}_s}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101}
\hat{\psi}_s(0) &= 0,\\[0.5ex]
\hat{\psi}_s(r_s) &= 1,\label{e102}\\[0.5ex]
\hat{\psi}_s(r\geq r_w) &= 0.\label{e103}
\end{align}

@@

\subsection{Solution in Presence of Resistive Wall}\label{resistive}
Suppose, now, that the wall at $r=r_w$ possesses a finite electrical resistivity, but is surrounded by a second perfectly-conducting wall
located at radius $r_c>r_w$.
The solution in the outer region can be written
\begin{equation}\label{e111}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r),
\end{equation}
where the real function $\hat{\psi}_s(r)$ is specified in the previous section, and the real function  $\hat{\psi}_w(r)$ is a solution of 
\begin{equation}\label{e100a}
\frac{d^2\hat{\psi}_w}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_w}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_w - \frac{J_z'\,\hat{\psi}_w}{r\,(1/q-n/m)}= 0
\end{equation}
that satisfies
\begin{align}\label{e101a}
\hat{\psi}_w(r\leq r_s) &= 0,\\[0.5ex]
\hat{\psi}_w(r_w) &= 1,\label{e102a}\\[0.5ex]
\hat{\psi}_w(r\geq r_c) &= 0.\label{e103a}
\end{align}
It is easily seen that
\begin{equation}\label{e117}
\hat{\psi}_w(r_w< r < r_c) = \frac{(r/r_c)^{-m} - (r/r_c)^m}{(r_w/r_c)^{-m} - (r_w/r_c)^m}.
\end{equation}

@@

The solution in the outer region can be written
\begin{equation}
\psi(r,t) = {\mit\Psi}_s(t)\,\hat{\psi}_s(r) + {\mit\Psi}_w(t)\,\hat{\psi}_w(r) + {\mit\Psi}_c\,\hat{\psi}_c(r),
\end{equation}
where the real functions $\hat{\psi}_s(r)$ and $\hat{\psi}_w(r)$ are specified in Sections~\ref{perfect} and \ref{resistive}, respectively. 
Moreover, 
the real function  $\hat{\psi}_c(r)$ is a solution of 
\begin{equation}\label{e100b}
\frac{d^2\hat{\psi}_c}{dr^2} + \frac{1}{r}\,\frac{d\hat{\psi}_c}{dr}-\frac{m^2}{r^2}\,\hat{\psi}_c=  0
\end{equation}
that satisfies
\begin{align}
\hat{\psi}_c(r\leq r_w) &= 0,\\[0.5ex]
\hat{\psi}_c(r_c) &= 1,\\[0.5ex]
\hat{\psi}_c(\infty) &= 0.
\end{align}
It is easily seen that
\begin{align}
\hat{\psi}_c(r\leq r_w) &= 0,\\[0.5ex]
\hat{\psi}_c(r_w<r \leq r_c) &= \frac{(r/r_w)^m-(r/r_w)^{-m}}{(r_c/r_w)^m - (r_c/r_w)^{-m}},\label{e142}\\[0.5ex]
\hat{\psi}_c(r> r_c) &=\left(\frac{r}{r_c}\right)^{-m}.\label{e143}
\end{align}

@@

Taking the scalar product of  $R_0\,{\bf e}_z$ with the plasma equation of motion, (\ref{e175}), and integrating around magnetic
flux-surfaces, similar arguments to those just employed for the ion poloidal flow allow us to write the zeroth-order ion toroidal angular velocity profile in the form
\begin{equation}
{\mit\Omega}_\varphi(r,t)= {\mit\Omega}_{\varphi\,0}(r) + {\mit\Delta\Omega}_\varphi(r,t),
\end{equation}
where ${\mit\Omega}_{\varphi\,0}(r)$ is the 
unperturbed (by the tearing mode) ion toroidal angular velocity profile, and ${\mit\Delta\Omega}_{\varphi}(r,t)$ is the  modification to  this  profile
induced by the toroidal electromagnetic torque that develops at the resonant surface. The perturbed ion toroidal angular
equation of motion is written
\begin{equation}\label{e185}
4\pi^2\,R_0^{\,3}\left[\rho\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s).
\end{equation}
Note that the ion parallel viscosity tensor does not give rise to damping of the toroidal flow profile. Furthermore, there is  no
neoclassical enhancement of the plasma toroidal ion inertia.

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In general, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp,s}\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf b}\right] = \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel$ and $\eta_\perp=1/\sigma_\perp$ are the parallel and perpendicular plasma
resistivities,  respectively. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$.

Next, we can add the electron and ion equations of motion, (\ref{e3.79b}) and (\ref{e3.80b}), again neglecting electron inertia and electron viscosity, 
to give the {\em plasma equation of motion}: 
\begin{equation}
n_0\,m_i\!\left[\frac{\partial {\bf V}}{\partial t} + ({\bf V}\cdot\nabla ){\bf V} + \frac{1}{1+\tau}\,({\bf V}_\ast\cdot\nabla){\bf V}_E\right]
+ \nabla p + \nabla\cdot\bpi_i - {\bf j}\times {\bf B}= 0.
\end{equation}
Here, $\nabla\cdot\bpi_i$ includes the contribution from the anomalous (i.e., enhanced due to the action of small-scale plasma turbulence) perpendicular viscosity tensor. The contribution from the gyroviscosity tensor,
which is the same size as the ion inertial terms, has already been incorporated into the equation (Hazeltine \& Meiss 1992). For the moment, we are neglecting the contribution of the parallel viscosity tensor.

Finally, we can combine the electron and ion energy evolution equations, (\ref{e3.79c}) and (\ref{e3.80c}), to obtain
a {\em plasma energy evolution equation}:
\begin{equation}
\frac{\partial p}{\partial t} + {\bf V}\cdot\nabla p + {\mit\Gamma}\,p\,\nabla\cdot{\bf V} + \frac{2}{3}\,\nabla\cdot {\bf q}=0, 
\label{e223}
\end{equation}
where ${\mit\Gamma}=5/3$. Here, we have neglected a number of  terms (e.g., the ohmic heating terms) because they
are, at least, factors $({\mit\Omega}_e\,\tau_e)^{-1}$ or  $({\mit\Omega}_i\,\tau_i)^{-1}$ smaller than $\nabla\cdot{\bf q}$. We have also neglected the viscous heating term because we
expect the dominant contribution, which comes from the parallel viscosity, to be much smaller than its nominal magnitude in a large aspect-ratio tokamak
plasma.

@@

It is convenient to adopt the following normalization scheme that renders all quantities in the drift-MHD equations dimensionless: 
$\hat{\nabla} = l\,\nabla$, $\hat{d}_i=d_i/l$, $\hat{t} = t/(l/V_A)$, $\hat{\bf B} = {\bf B}/B_z$, $\hat{\bf E} = {\bf E}/(B_z\,V_A)$,
$\hat{\bf j} = {\bf j}/(B_z/\mu_0\,l)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,V_A\,l)$, $\hat{\bf V} = {\bf V}/V_A$,
$\hat{{\bf V}}_{\ast,i}= {\bf V}_{\ast,i}/V_A$, $\hat{V}_{\parallel\,i}= V_{\parallel\,i}/V_A$, $\hat{p} = p/(B_z^{\,2}/\mu_0)$, $\hat{p}_0 = p_0/(B_z^{\,2}/\mu_0)$, $\hat{\bpi}_{i} = \bpi_{i}/(B_z^{\,2}/\mu_0)$,
$\hat{\bf q} = {\bf q} /(B_z^{\,2}\,V_A/\mu_0)$. Equations~(\ref{e220})--(\ref{e226}) yield
\begin{align}\label{e227}
\hat{\bf E} + \hat{\bf V}\times \hat{\bf B} + \hat{d}_i\!\left[\hat{\nabla} \hat{p} + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\hat{\nabla} \hat{p})\,{\bf b} + \hat{\nabla}\cdot\hat{\bpi}_e-\hat{\bf j}\times\hat{\bf b}\right] &= \hat{\eta}_\parallel\,\hat{\bf j}_\parallel+ \hat{\eta}_\perp\,\hat{\bf j}_\perp,\\[0.5ex]
\frac{\partial \hat{\bf V}}{\partial \hat{t}} 
+ (\hat{\bf V}\cdot\hat{\nabla} )\hat{\bf V} + 
\frac{1}{1+\tau}\,(\hat{\bf V}_\ast\cdot\hat{\nabla})\hat{\bf V}_E
+ \hat{\nabla} \hat{p} + \hat{\nabla}\cdot\hat{\bpi}_i -
 \hat{\bf j}\times \hat{\bf B} &= 0,
 \\[0.5ex]
\frac{\partial \hat{p}}{\partial \hat{t}} + \hat{\bf V}\cdot\hat{\nabla} \hat{p} + {\mit\Gamma}\,\hat{p}\,\hat{\nabla}\cdot\hat{\bf V} + \frac{2}{3}\,\hat{\nabla}\cdot \hat{\bf q}&= 0,\label{e229}
\end{align}
and
\begin{align}
\hat{\nabla}\cdot\hat{\bf B} &= 0,\label{e230}\\[0.5ex]
\hat{\nabla}\times \hat{\bf E} &= - \frac{\partial \hat{\bf B}}{\partial \hat{t}},\label{e231}\\[0.5ex]
\hat{\bf j} &= \hat{\nabla} \times \hat{\bf B}.\label{e232}
\end{align}
Finally,
\begin{align}
\hat{\bf V}_E &= \hat{\bf E}\times {\bf b},\\[0.5ex]
\hat{\bf V} &= \hat{\bf V}_E + \hat{V}_{\parallel\,i}\,{\bf b},\\[0.5ex]
\hat{\bf V}_\ast &= \hat{d}_i\,{\bf b}\times \hat{\nabla} \hat{p},\\[0.5ex]
\hat{\bf V}_i &= \hat{\bf V} + \frac{\hat{\bf V}_{\ast\,i}}{1+\tau}.
\end{align}

\subsection{Reduction Process} 
All variables in the resonant layer are assumed to be functions of $\hat{x}=(r-r_s)/l$ and $\zeta=m\,\theta-n\,\varphi$ only. 
Let ${\bf n} = (0,\,\epsilon/q_s,\,1)$, where $\epsilon=r/R_0$ and $q_s=m/n$. It follows that ${\bf n}\cdot\nabla A=0$ for any $A(\hat{x},\zeta)$. 
Let us write (Fitzpatrick \& Waelbroeck 2009)
\begin{equation}\label{e238}
\hat{\bf B} \simeq (1+ \delta b)\,{\bf n} + \hat{\nabla}\hat{\psi}\times {\bf n},
\end{equation}
which automatically satisfies Equation~(\ref{e230}). (Recall that we are assuming that $|r-r_s|\ll r_s$ inside the resonant layer.) Note that $\hat{\psi}=\psi/r_s$, where $\psi$ is defined in Equation~(\ref{e93}).
 Here, $\delta b(\hat{x},\zeta)$ and $\hat{\psi}(\hat{x},\zeta)$ are assumed to be small
compared to unity, which merely indicates that the perturbed magnetic field due to the tearing mode is small compared to the
equilibrium magnetic field.  We can write
\begin{equation}
\hat{\bf E} \simeq \left(\hat{E}_\parallel- \frac{\partial \hat{\psi}}{\partial \hat{t}}\right)\!{\bf n} + \hat{\nabla}\!\left(\frac{\partial\chi}{\partial \hat{t}}\right)\times {\bf n} + \hat{\nabla}\phi,
\end{equation}
which automatically satisfies Equation~(\ref{e231}) provided that
\begin{equation}
\hat{\nabla}^2\chi =\delta b.
\end{equation}
Here, $\hat{E}_\parallel = E_\parallel/(B_z\,V_A)$, where $E_\parallel$ represents the constant inductive electric field that drives the equilibrium plasma current in the vicinity of the resonant surface. Moreover, $\partial/\partial \hat{t}$ and $\phi(\hat{x},\zeta)$
are 
assumed to be first order in small quantities, whereas $\hat{E}_\parallel$ is assumed to be second order. 
It follows that
\begin{align}
\hat{\bf V}_E &\simeq \hat{\nabla}\phi\times {\bf n},\\[0.5ex]
\hat{\bf V} &\simeq \hat{V}_{\parallel\,i}\,{\bf n}+  \hat{\nabla}\phi\times {\bf n}+\hat{\nabla}{\mit\Xi},
\end{align}
where $\hat{V}_{\parallel\,i}(\hat{x},\zeta)$ is assumed to be first order in small quantities, whereas ${\mit\Xi}(\hat{x},\zeta)$ is assumed to be
second order. 
We can write
\begin{equation}
\hat{p}= \frac{\beta}{{\mit\Gamma}} + \delta p,
\end{equation}
where 
\begin{equation}\label{e244}
\beta= {\mit\Gamma}\,\hat{p}_0= \frac{{\mit\Gamma}\,p_0\,\mu_0}{B_z^{\,2}}.
\end{equation}
 Here, $\delta p(\hat{x},\zeta)$ is assumed to be first order in small
quantities. 
It follows that
\begin{align}
\hat{\bf V}_\ast &\simeq \hat{d}_i\,{\bf n}\times \hat{\nabla} \delta p,\\[0.5ex]
\hat{\bf V}_i&\simeq  \hat{V}_{\parallel\,i} \,{\bf n}+ \hat{\nabla}\!\left(\phi - \frac{\hat{d}_i\,\delta p}{1+\tau}\right)\times {\bf n}.
\end{align}
